Accesible, realistic genome simulation with selection using \stdpopsim

### 2025-02-13 analysis2_manuscript.tex

    % goals
    In this manuscript, we provide an overview of a major new addition
    to \stdpopsim---the inclusion of models of selection.
    We begin by describing the models of selection that we have implemented,
    as well as the parameter choices that are available to the user.
    This includes models of background selection, selective sweeps, and
    models of the distribution of fitness effects (DFE).
    Further we describe how these models can be combined with genomic
    annotations that are specific to a given species available
    through the \stdpopsim API, and combined with a parameterized model of
    demography and recombination to provide a realistic simulation of
    genetic variation in a population.
    We then provide a series of examples of how these models can be used
    to benchmark and compare the performance of different methods for
    inferring demographic history, the targets of recent positive selection,
    and the distribution of fitness effects from genetic data.

@@

\section*{Example Applications with Selection}
    \label{applications}
    In this section, we present a series of examples of how the new models of selection
    in \stdpopsim can be used to benchmark different
    methods for population genetic inference.
    We focus on three main areas: demographic inference in the 
    single and multi-population settings, inference of the DFE,
    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the pairwise sequentially Markovian coalescent
    (PSMC) \citep{li2011inference}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}. 
    While these methods perform well under neutral evolution, because of its effects on
    patterns of genetic variation, natural selection on linked sites can bias estimates of $N_e(t)$
    away from true census population sizes simply because selection increases the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}.

    \igcomment{Questions about methods here:
    (1) Which DFE model was used here? The previous version of the text cited Huber, but the production config says Kim;
    (2) Was DFE applied to all exonic sites or nonsynonymous? (I think that all exonic sites, but if non-syn, we need to clarify);
    (3) We mention 3 reps, but this is not used in the results (see also comment in next para)
    }
    Using \stdpopsim, we can easily examine the effect of selection on the effective
    population sizes inferred by different methods.
    For this purpose, we simulated human genomes with and without natural selection.
    In both scenarios, we ran three replicate simulations
    under the out of Africa (OOA) demographic model
    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second}.
    Simulations with selection were implemented by modeling purifying selection on exons,
    defined from the HAVANA group release 104 for the human genome,
    using a DFE inferred by \cite{kim2017inference}.
    Each of the six simulated datasets contained 300 genomes, 100 for each geographic group: CEU, CHB, and YRI.
    All simulations were implemented using a relatively short python script (as illustrated in Figure 1B).
    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
    % IG: added a note about this above.
    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.

    \igcomment{Questions about inference and summary:
    (1) In simulation with selection, do we infer Ne from all sites (exons and non-exons)?
        This is what I understand from the code, but if not, we should clarify.
    (2) The effect of selection on inference is difficult to glean from Fig 2.
        The main thing I see there is that the influence is very small.
        Maybe we should have a separate plot showing each method in a diff panel with inferences from all 6 datasets?
        This might allow us to see differences between the two scenarios (with/without selection).  
    }
    %IG: I commented this out because it repeats some of the info above and we now start by directly
    %discussing the results rather than describing the figure
    %In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
    %from data simulated under the five-population model of human out-of-Africa demography
    %from Ragsdale \textit{et al.} (2019)
    %with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
    %The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
    %while the right column shows the results
    %from simulations with deleterious mutations occurring in exons.
     % IG: commented this out, because these details are better left for the fig caption
    %In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
    %in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
    %from a model without selection.
    We see that all methods generally perform reasonably well in the absence of selection (Figure \ref{fig:1pop-human-demography}; left).
    The only exception to this is that GONE seems to overestimate $N_e(t)$ from the CHB genomes,
    %IG: I commented this out because this seems like a very minor blip
    %and to a lesser extent in the CEU sample.
    presumably due to patterns of linkage disequilibrium that result from the combination
    of bottleneck and growth in the populations ancestral to these samples.
    %
    \igcomment{I don't really see evidence for the observations listed below in Fig 2. See my suggestion above to enhance the comparison.}
    In the presence of selection, different methods behave differently. 
    For instance stairwayplot, which uses the SFS to infer $N_e(t)$, 
    is reasonably close to the true $N_e(t)$ both with and without selection,
    although it seems to infer significant population size changes that are not part of the true history.
    Likewise \smcpp seems to be robust to the presence of selection,
    but the timing of population size changes is slightly inaccurate. 
    \msmc seems to be the most affected by the presence of selection, 
    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$ particularly in the very recent past,
    although estimates from \msmc have quite a bit of added noise even in the absence of selection.

    % TODO: legend here needs to be changed. The dotted line indicates no selection, the solid selection

    \begin{figure}[t]
        \centering
        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \caption{
        \label{fig:1pop-human-demography}
        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        \igcomment{This is a nice figure but it's difficult to see if there is any subtle influence of selection from it.
        We should think of one multi-panel main figure to summarize different aspects of the $N_e(t)$ analysis, and push the rest to the supplement.
        Specifically for this figure, I'm not sure we need to use different line types for left and right since the plots are separated.}  
        }
    \end{figure}
    %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...
    %IG: this seems to have been taken care of

    We note that while selection in these simulations directly acts on a small fraction of sites in the analyzed genomes,
    it has a widespread influence on genetic diversity through background selection \citep[e.g.][]{charlesworth1993effect}.
    Our results here suggest that the combined effect of direct selection and background selection does not considerably
    influence infrerence of $N_e(t)$ for human populations under reasonable assumptions about their evolutionary history.
    %
    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and since no genetic map is currently available, we assumed a constant recombination rate across the genome.
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly \igcomment{Maybe cite Morin et al. 2010 for assembly and annotations?}.
    \igcomment{I added this about the h-s relationship. Please check!}
    In this DFE model, we implemented a relationship between the dominance coefficient ($h$) and the selection coefficient ($s$)
    as follows: $s<-0.1 \Rightarrow h=0$; $s\in [-0.1,-0.01) \Rightarrow h=0.01$; $s\in [-0.01,-0.001) \Rightarrow h=0.1$;
    and $s\geq -0.001 \Rightarrow h=0.4$.
    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically} and is easily
    implemented in \stdpopsim's \texttt{DFE} class.
    Using these simulations, we see that all methods infer $N_e(t)$ reasonably well with or without selection
    (Figure \ref{fig:1pop-vaquita-demography}), although they do not accurately identify the boundary time
    between the two epochs.
    % MFR: Should we say somewhere that people use to expect from BGS? That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
    % IG: we can address this in the discussion.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
    \caption{
    \label{fig:1pop-vaquita-demography}
    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
    The left panel shows estimates of $N_e(t)$ from simulations
    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
    \igcomment{see my comment on fig 2. We should decide what part of this stays in the main figure and what ends up in the supp.}  
    }
\end{figure}

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
    the performance of different methods for inferring the DFE. 
    %IG: I don't think we need this intro.
    %We performed simulations using the 
    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
    %
    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
     %https://academic.oup.com/genetics/article/207/3/1103/5930676
    Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
    effective population size $N_e$, and different methods assume different scaling.
    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
    %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
    For comparison, here we have normalized all inference results to the convention of SLiM.
    \igcomment{Can we summarize this in a table or in a short methods section?
    For each method, we could specify the scaling we apply to compare its inferred distribution to the one used in simulation.}

@@

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons.
        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; left) and shape parameter (right).
        The DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli. The dashed horizontal lines represents the simulated values.
        \igcomment{Does the boxplot represent the three simulation reps? Is each method run several times?}
}
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
    We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
    with selection acting on exons (same data used for Figure \ref{fig:1pop-human-demography}; right).
    In this setting, DFE is inferred separately for each of the three simulated populations
    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as \stdpopsim's models currently do not
    contain outgroup populations or species that can be used to estimate divergence.
    %IG: no need to map figures. Just describe results and cite figs
    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
    The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
    Population history appears to influence different methods in slightly different ways.
    For instance, \dadicli and \grapes produce more accurate estimates from the YRI genomes 
    (relative to the other two populations), whereas \polydfe
    produces the most accurate estimate from the CHB genomes.
    To factor out the influence of demographic changes, we also analyzed genomes generated using similar
    simulations, but with a simpler demographic model with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
    In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
    the other methods do not. 
    \grapes exhibits similar biases here as in the simulations with more complex demography,
    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
    with overestimated shape parameter and slightly underestimated mean. 
\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things. That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models. I would also do this for the demogrpahic models as well as the DFE models}
\igcomment{I suspect that this way of comparing inferred distributions to simulated ones might be a bit misleading. Wouldn't it be better to use something like what we do in fig 5C?
I predict that if we do this, we'll get a reasonable match for \dadicli}

@@

    %IG: describe experiment and cite figure later
    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome,
    as generated for the analysis shown in Fig. \ref{fig:1pop-vaquita-demography} (right).
    Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
    and a relationship between the selection and dominance coefficients.
    In this setting, all methods perform uniformly worse than in the human genome simulations,
    with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
    Similar patterns are observed when re-running the simulations with a simpler demographic model
    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection (see above). 
    To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
    simulated distribution of $2hs$, which would have been equal to $s$ under an additive model (Figure \ref{fig:vaquita-dfe}C).
    Reassuringly, all three methods infer distributions that appear to fit the main mode of the simulated distribution of $2hs$.
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. As stated above, simulations did or did not include a model 
    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
    Three replicates of whole-genome simulations were performed for each species, with 
    a sample size of 100 diploid individuals. Demographic models are as stated above, each
    using the exon annotations from the human and vaquita porpoise genomes respectively.
    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations. \kelcomment{I'm confused about the 100 diploid individuals samples vs. the 10 diploid individuals sampled. Did some analyses use 100 and others use 10? That's fine, but maybe we can specify that more clearly...}

    \subsection*{Demographic inference}
    We assessed the accuracy of demographic inferences from
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
    Each of these methods was run using the same set of simulations in replicate.
    In each case, the non-recombining portion
    of the genome was masked to remove any bias due to recombination rate % DRS: might need more detail here
    variation. Comparisons between inferences made with and without selection were 
    of particular interest. For \msmc, we used a random sample of $N=6$ genomes with 20
    iterations of the EM algorithm. For \gone, we set max\_snps to 500000,
    the number of generations to 2000, and the number of bins to 400, all representing default settings.

>>>>

    % goals
    In this manuscript, we provide an overview of a major new addition
    to \stdpopsim---the inclusion of models of selection.
    We begin by describing the models of selection that we have implemented,
    as well as the parameter choices that are available to the user.
    This includes models of background selection, selective sweeps, and
    models of the distribution of fitness effects (DFE).
    Further we describe how these models can be combined with genomic
    annotations that are specific to a given species available
    through the \stdpopsim API, and combined with a parameterized model of
    demography and recombination to provide a realistic simulation of
    genetic variation in a population.
    We then provide a series of examples of how these models can be used
    to benchmark and compare the performance of different methods for
    inferring demographic history, the distribution of fitness effects,
    and the targets of recent positive selection, from genetic data.

@@

%IG: commented out this intro paragraph. Better to dive right in.
%\section*{Example Applications with Selection}
%    \label{applications}
%    In this section, we present a series of examples of how the new models of selection
%    in \stdpopsim can be used to benchmark different
%    methods for population genetic inference.
%    We focus on three main areas: demographic inference in the 
%    single and multi-population settings, inference of the DFE,
%    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the sequentially Markovian coalescent
    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}.
    Since all of these methods assume neutral evolution, they are typically applied to genomic
    segments that are not expected to be directly affected by selection (e.g., by masking exons
    and conserved elements).
    However, natural selection acting on linked sites (that are masked out of the analysis)
    can bias estimates of $N_e(t)$
    away from true census population sizes by increasing the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
    which is commonly referred to as \emph{background selection}, 
    on different methods for inferring $N_e(t)$.

    For this purpose, we simulated human genomes with and without natural selection.
    In both scenarios, we ran three replicate simulations
    under the out of Africa (OOA) demographic model of
    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
    Simulations with selection were implemented by modeling purifying selection on exons
    % IG: moved details to methods
    %, defined from the HAVANA group release 104 for the human genome,
    using a DFE inferred by \cite{kim2017inference}.
    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.

    \begin{figure}[b!]
        \centering
        %IG: NEED TO LINK NEW FIG HERE
        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
        \caption{
        \label{fig:human-demography}
        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. 
        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
        Columns correspond to the four inference methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
        and the three datasets simulated with the background effects of purifying selection on exons (orange),
        alongside the true values of $N_e$ used in simulation (black).
        }
    \end{figure}

    \igcomment{Rewrote this paragraph based on new figs.}
    %
    All four inference methods appear to infer similar $N_e(t)$ trajectories from
    genomes simulated without selection (Figure \ref{fig:human-demography}; blue) and
    genomes simulated with selection (Figure \ref{fig:human-demography}; orange).
    Thus, background selection does not appear to notably bias the estimates
    produced by these methods.
    As noted in previous studies that compared demography inference methods
    on neutral simulations \citep{adrion2020community},
    we see that each method has its strengths and weaknesses.
    Methods based on the sequenial Markovian coalescent (\smcpp and \msmc)
    appear to overal produce the most accurate trajectories of $N_e(t)$
    in this setting, although they tend to under-estimate the population
    bottleneck associated with the population split.
    On the other hand, \stairway, which utilizes the site frequency spectrum,
    produces more noisy estimates,
    but it does not appear to under-estimate the population bottleneck.
    The inference of \gone, which utilizes identity by descent,
    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
    explaining its overall noisy estimates.

%IG: commented old version of figure here
%    \begin{figure}[t]
%        \centering
%        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
%        \caption{
%        \label{fig:1pop-human-demography}
%        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
%        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
%        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
%        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
%        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
%        }
%    \end{figure}

    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and a constant recombination rate across the genome (see \textbf{Methods}).
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly.
    This DFE model implements a relationship between selection and dominance 
    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
    additive ($h=0.4$; see \textbf{Methods}).
    Here, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these difference, we see that background selection does not
    considerably influence the inference of any of the four methods also for this species,
    which has a very different evolutionary history than humans.

    % MFR: Should we say somewhere that people use to expect from BGS? That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
    % IG: we can address this in the discussion, or in the revised text above.

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
%    \caption{
%    \label{fig:1pop-vaquita-demography}
%    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
%    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
%    The left panel shows estimates of $N_e(t)$ from simulations
%    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
%    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
%    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
%
%    }
%\end{figure}

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
    the performance of different methods for inferring the DFE. 
    %IG: I don't think we need this intro.
    %We performed simulations using the 
    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
    %
    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
     %https://academic.oup.com/genetics/article/207/3/1103/5930676
    Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
    effective population size $N_e$, and different methods assume different scaling.
    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
    %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
    For comparison, here we have normalized all inference results to the convention of SLiM.
    \igcomment{We should add a detailed description in the Methods of what we do to scale the output of each DFE-inference method.}

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
        \igcomment{NEED TO UPDATE FIG WITH NEW VERSION}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
    We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
    with selection acting on exons (see \textbf{Methods}).
    In this setting, DFE is inferred separately for each of the three simulated populations
    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as \stdpopsim's models currently do not
    contain outgroup populations or species that can be used to estimate divergence.
    %IG: no need to map figures. Just describe results and cite figs
    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
    The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters (Figure \ref{fig:homsap-dfe.ooa}A-B).
    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$; Figure \ref{fig:homsap-dfe.ooa}C),
    while \grapes under-estimates the fraction of mutations with $s<0.01$ and considerably over-estimates the fraction of
    mutations under weak selection ($s\geq-0.001$).
    Population history appears to influence different methods in slightly different ways.
    For instance, \dadicli and \grapes produce more accurate estimates from the YRI genomes 
    (relative to the other two populations), whereas \polydfe
    produces the most accurate estimate from the CHB genomes.
    To factor out the influence of demographic changes, we also analyzed genomes generated using similar
    simulations, but with a simpler demographic model with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
    In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
    the other methods do not. 
    \grapes exhibits similar biases here as in the simulations with more complex demography,
    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
    with overestimated shape parameter and slightly underestimated mean. 
\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things. That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models. I would also do this for the demogrpahic models as well as the DFE models}
\igcomment{I think that this is better conveyed now in the added histograms for $s$ in panel C}

@@

    %IG: describe experiment and cite figure later
    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
    Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
    and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
    In this setting, all methods perform uniformly worse than in the human genome simulations,
    with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
    Similar patterns are observed when re-running the simulations with a simpler demographic model
    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection. 
    To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
    simulated distribution of $2hs$, which would have been equal to $s$ under an additive model (Figure \ref{fig:vaquita-dfe}C).
    Reassuringly, all three methods infer distributions that appear to fit the main mode of the simulated distribution of $2hs$.
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}

@@

    \subsection*{Simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework.
    All simulations were executed using the SLiM engine with a burn-in of $2N$ generations
    (where $N$ is the initial population size), and a scaling factor of 2 (to reduce running time).
    Each simulation was implemented using a relatively short python script, similar to the one shown in Figure 1B
    (see \textbf{Data availablitiy} below).
    \igcomment{We should add a link to the analysis2 repo in data/code availablity section}
    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
    % IG: added a note about this above.

    \subsubsection*{Simulations of complete human genomes}
    Simulation of human genomes used a genetic map from the HapMap Project
    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
    the out of Africa (OOA) demographic model with archaic admixture from
    \cite{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
    Each simulation generate 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
    with each genome consisting of 22 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
    Each of these simulations generated 100 diploid genomes.
